---
title: "02-Data_Visualization_Table_FAIR_Retrospective"
format: html
editor: visual
---

Read in data
Please run 01-Data_Cleaning_FAIR_Retrospective first
```{r}
library(readr)
library(tidyverse)
alex_doi_new1 <- read_csv("output_data/alex_doi_new1.csv")
```

Simple Summary Chart for the Numbers of DOI in the Domain, Field, and Subfield
```{r}
# Summary table grouped by the unique DOI
domain_summary <- alex_doi_new1 %>%
  filter(Domain !="") %>% 
  group_by(Domain) %>%
  summarise(
    unique_dois_count = n_distinct(doi, na.rm = TRUE),
    unique_field_count = n_distinct(Field, na.rm = TRUE),
    unique_subfield_count = n_distinct(Subfield, na.rm = TRUE)
  )

# Print the resulting table
print(domain_summary)
```

```{r}
# generate a table with count of the top Field
# Read the Google AI summary on how to add numbers for top field on 8/31 
field_table <- alex_doi_new1 %>%
  filter(!is.na(doi)) %>%
  mutate(
    Domain = coalesce(Domain, "Unclassified"),
    Field  = coalesce(Field,  "Unclassified")
  ) %>%
  group_by(Domain, Field) %>%
  summarise(doi_count = n_distinct(doi), .groups = "drop")

field_table <- field_table %>%
  group_by(Domain) %>%
  mutate(
    is_top_field = doi_count == max(doi_count),
    label_text = if_else(
      is_top_field,
      paste0(Field, "\n(n = ", doi_count, ")"),
      Field
    )
  ) %>%
  ungroup()
```

Treemap 
```{r}
okabe_ito <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2") 

library(showtext)
font_add(family = "Arial", regular = "Arial.ttf")
showtext_auto()

ggplot(field_table, aes(area = doi_count, fill = Domain, label = label_text, subgroup = Domain)) +
  geom_treemap() +
  geom_treemap_text(colour = "black", place = "center", reflow = TRUE, size = 14,
                     family = "Arial", min.size =5) +
  geom_treemap_subgroup_border(colour = "darkgrey", size = 0.5) +
  scale_fill_manual(values = okabe_ito) +
  labs(title = "Distribution of Publications by Domain and Field") 
```

#### Table of countries ####
# list of countries
Logic for the countries count. Generated with the assistance of Google ai mode on July 1 2026.
```{r}
DATE_MIN   <- as.Date("2016-03-15")
DATE_MAX   <- as.Date("2026-03-16")

code_col <- intersect(c("code", "iso2", "iso2c", "country_code"), names(country_dict))
name_col <- intersect(c("name", "country_name", "country", "label"),  names(country_dict))

if (length(code_col) == 0) stop("country_dict must have a code column.")
country_dict_std <- if (length(name_col) == 0) {
  country_dict %>% rename(code = !!code_col[1]) %>% mutate(name = code)
} else {
  country_dict %>% rename(code = !!code_col[1], name = !!name_col[1])
}

country_codes <- names(alex_doi_new_wide)[str_detect(names(alex_doi_new_wide), "^[A-Z]{2}$")]
```


```{r}
df_long <- alex_doi_new_wide %>%
  filter(publication_date >= DATE_MIN, publication_date <= DATE_MAX) %>%
  mutate(publication_year = as.integer(publication_year)) %>%
  filter(!is.na(publication_year), publication_year > 2015) %>%
  select(publication_year, doi_clean, all_of(country_codes)) %>%
  pivot_longer(cols = all_of(country_codes),
               names_to = "country_code", values_to = "value") %>%
  filter(!is.na(value), value > 0) %>%
  left_join(country_dict_std, by = c("country_code" = "code")) %>%
  mutate(country_name = coalesce(name, country_code)) %>%
  select(publication_year, doi_clean, country_code, country_name, value)

summary_full <- df_long %>%
  group_by(publication_year, country_code, country_name) %>%
  summarise(count = n_distinct(doi_clean), .groups = "drop") %>%
  group_by(publication_year) %>%
  mutate(year_total = sum(count)) %>%
  ungroup()

all_countries <- sort(unique(summary_full$country_name))

```

# Table for Countries count by unique DOI's
```{r}
countries_doi_count <- df_long  %>%
  group_by(country_name) %>%
  summarise(unique_items = n_distinct(doi_clean), .groups = 'drop')
```


#### Finding the Subfields not represented ####
We have coverage in all domains and fields. and 210 out of 252 subfields in OpenAlex.
claude.ai prompt for Sonnet 5 Medium on Aug. 24, 2026 - I asked create a API pull in OpenAlex using R to get subfields data. There are 252 subfields. The Subfields endpoint is at `api.openalex.org/subfields`.
What are they?
I modified the response code to...
```{r}
library(httr2)
library(dplyr)
library(purrr)
library(tibble)

# Pull all subfields via cursor pagination (per_page max = 200,
# and there are 252 subfields, so this will take 2 pages)
get_openalex_subfields <- function(mailto = "jungmee.park@wichita.edu") {
  
  base_url <- "https://api.openalex.org/subfields"
  cursor   <- "*"
  all_results <- list()
  
  repeat {
    resp <- request(base_url) %>%
      req_url_query(
        per_page = 200,
        cursor   = cursor,
        mailto   = mailto          
      ) %>%
      req_perform()
    
    parsed <- resp_body_json(resp, simplifyVector = FALSE)
    
    all_results <- c(all_results, parsed$results)
    
    cursor <- parsed$meta$next_cursor
    if (is.null(cursor) || length(parsed$results) == 0) break
  }
  
  all_results
}

subfields_raw <- get_openalex_subfields(mailto = "jungmee.park@wichita.edu")
length(subfields_raw)   # should be 252

# Flatten into a tidy data frame: subfield -> its parent field -> its parent domain
subfields_df <- map_dfr(subfields_raw, function(x) {
  tibble(
    subfield_id    = x$id,
    subfield_name  = x$display_name,
    field_id       = x$field$id,
    field_name     = x$field$display_name,
    domain_id      = x$domain$id,
    domain_name    = x$domain$display_name,
    works_count    = x$works_count %||% NA_integer_,
    cited_by_count = x$cited_by_count %||% NA_integer_
  )
})

# Strip the "https://openalex.org/" prefix off the IDs so you're left with the
# bare numeric/alpha IDs, matching the style you used earlier (e.g. "2718")
subfields_df <- subfields_df %>%
  mutate(
    subfield_id = sub("https://openalex.org/subfields/", "", subfield_id),
    field_id    = sub("https://openalex.org/fields/",    "", field_id),
    domain_id   = sub("https://openalex.org/domains/",   "", domain_id)
  ) %>%
  arrange(domain_name, field_name, subfield_name)


```
Google prompt "can you compare two variables in two different datasets" on 8/24/2026

Google ai prompt. dataframe A and dataframe b compare the different responses in the same columns (column names are slightly different) in R (8/24/2026)
# response code
# # Example if columns are shuffled
# library(dplyr)
# 
# df_B <- df_B %>% 
#   rename(col1 = slight_diff_name1, col2 = slight_diff_name2) %>% 
#   select(colnames(df_A))

Use library(waldo) to find missing subfields
```{r}
subfields_df <- subfields_df |> 
  mutate(Subfield = subfield_name)

library(waldo)

subfields_df_A <- subfields_df$Subfield
subfields_df_B <- alex_doi_new1$Subfield

# Elements in A that are completely missing from B
unshown_subfields <- setdiff(subfields_df_A, subfields_df_B)
```


response modified to
```{r}
unknown_subfields <- subfields_df |> 
  filter(Subfield %in% unshown_subfields)

```



