---
title: "Open_alex_static_visuals"
format: html
editor: visual
---

Sunburst chart August 17, 2026 modified after using a new dataset that merged from open alex, downloaded August 7, 2026

create tables with this information

```{r}
library(DT)

journal_hierarchy <- alex_doi_new1 %>%
  filter(!is.na(Domain), 
         # !is.na(field), !is.na(subfield),
         !is.na(publication_year),
         !is.na(primary_location.source.display_name)) %>%
  distinct(publication_year, Domain, Field, Subfield,
           primary_location.source.display_name) %>%
  count(publication_year, Domain, Field, Subfield,
        name = "n_unique_journals") %>%
  arrange(publication_year, Domain, Field, Subfield)

datatable(
  journal_hierarchy,
  filter   = "top",           # per-column filter boxes
  rownames = FALSE,
  colnames = c("Year", "Domain", "Field", "Subfield", "Unique Journals"),
  options  = list(
    pageLength  = 20,
    scrollX     = TRUE,
    autoWidth   = TRUE,
    columnDefs  = list(list(className = "dt-center", targets = c(0, 4)))
  )
)

```
clean up the names of domain, field, subfield
```{r}
# Summary table grouped by the unique DOI
domain_summary <- alex_doi_new1 %>%
  filter(Domain !="") %>% 
  group_by(Domain) %>%
  summarise(
    unique_dois_count = n_distinct(doi, na.rm = TRUE),
    unique_field_count = n_distinct(Field, na.rm = TRUE),
    unique_subfield_count = n_distinct(Subfield, na.rm = TRUE)
  )


# Print the resulting table
print(domain_summary)
```

```{r}
# Create a flattened hierarchical table
sunburst_table <- alex_doi_new1 %>%
  group_by(Domain, Field, Subfield) %>%
  summarise(
    doi_count = n_distinct(doi, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Sort from top tier down to make the hierarchy easy to read
  arrange(Domain, Field, Subfield)

print(sunburst_table)
```

```{r}
# Create a clean summary table
domain_counts <- alex_doi_new1 %>%
  group_by(Domain) %>%
  summarise(
    doi_count = n_distinct(doi, na.rm = TRUE),
    .groups = "drop",
    field_count = n_distinct(Field),  # Counts unique schools
    subfield_count = n()                                   # Counts total sports rows
  )

grid_counts <- alex_doi_new1 %>%
  count(Field, Subfield) %>%
  pivot_wider(names_from = Subfield, values_from = n, values_fill = 0)

print(grid_counts)
```

# attempt a dendogram
```{r}
# Libraries
library(ggraph)
library(igraph)
library(tidyverse)
theme_set(theme_void())


```


```{r}
# Plotly Sunburst visual
library(plotly)
library(dplyr)

sunburst_data <- alex_doi_new1 %>%
  filter(!is.na(Domain), 
         # !is.na(field), !is.na(subfield),
         !is.na(primary_location.source.display_name)) %>%
  distinct(Domain, Field, 
           # Subfield,
           primary_location.source.display_name) %>%
  count(Domain, Field, name = "n_unique_journals") # took out Subfield,

# Build ids/labels/parents for plotly sunburst
domains  <- sunburst_data %>%
  group_by(Domain) %>%
  summarise(n = sum(n_unique_journals)) %>%
  transmute(ids = Domain, labels = Domain, 
            parents = "", values = n)

fields   <- sunburst_data %>%
  group_by(Domain, Field) %>%
  summarise(n = sum(n_unique_journals), .groups = "drop") %>%
  transmute(ids = paste(Domain, Field, sep = " - "),
            labels = Field, parents = Domain, values = n)

# subfields <- sunburst_data %>%
#   transmute(ids     = paste(Domain, Field, Subfield, sep = " - "),
#             labels  = Subfield,
#             parents = paste(Domain, Field, sep = " - "),
#             values  = n_unique_journals)

sb <- bind_rows(domains, fields) # , subfields

plot_ly(
  sb,
  ids     = ~ids,
  labels  = ~labels,
  parents = ~parents,
  values  = ~values,
  type    = "sunburst",
  branchvalues = "remainder"
  # branchvalues = "total"
)
```

```{r}
table(alex_doi_new1$Domain)
```

make a publishable table asked google ai with generic code august 20, 2026
```{r}
library(gtsummary)
library(gt)

# Generate the table including unclassified (NA) values
# (Replacing your original command with an explicit useNA argument)
domain_table <- table(alex_doi_new1$Domain, useNA = "ifany")

# 2. Convert the table into a clean data frame for formatting
table_df <- as.data.frame(domain_table)
colnames(table_df) <- c("Domain", "Frequency")

# 3. Clean up the "unclassified" label if it appears as <NA>
table_df$Domain <- as.character(table_df$Domain)
table_df$Domain[is.na(table_df$Domain)] <- "Unclassified"

domain_freq_table <- table_df|> 
  arrange(desc(Frequency)) |> 
  gt() |> 
  cols_align(align = "left", columns = Domain) %>%
  cols_align(align = "right", columns = Frequency) %>%
    # wt  = "Weight (1000 lbs)"
  tab_header(
    title = "Research Domains and the Unique DOI's",
    subtitle = "Base on OpenAlex"
  ) 

# 4. Build and style the professional table
domain_freq_table <- table_df %>%
  gt() %>%
  # Align text left and numbers right
  cols_align(align = "left", columns = Domain) %>%
  cols_align(align = "right", columns = Frequency) %>%
  cols_label(
    Domain= "Domain",
    Frequency  = "Unique DOI's",
  ) 
  # # Bold the column headers
  # tab_style(
  #   style = cell_text(weight = "bold"),
  #   locations = cells_column_labels()
  # )

# Print the table
domain_freq_table
# countries_doi_count <- df_long  %>%
#   group_by(country_name) %>%
#   summarise(unique_items = n_distinct(doi_clean), .groups = 'drop')
# 
# most_active_countries <- countries_doi_count %>% 
#   filter(unique_items >= 1000)
# # Create a publication-ready table 1
# 
# most_active_countries_table <- most_active_countries|> 
#   arrange(desc(unique_items)) |> 
#   gt() |> 
#   cols_label(
#     country_name = "Country of Author(s)",
#     unique_items  = "Unique DOI's",
#     # wt  = "Weight (1000 lbs)"
#   ) |> 
#   tab_header(
#     title = "Top 10 Most Active Countries of Authorship",
#     subtitle = "Base on OpenAlex"
#   ) 
gtsave(domain_freq_table, filename = "visual_output/domain_freq_table.docx")
```



```{r}
sunburst_data <- alex_doi_new1 %>%
  filter(!is.na(Domain), 
         # !is.na(field), !is.na(subfield),
         !is.na(primary_location.source.display_name)) %>%
  distinct(Domain, Field, Subfield,
           primary_location.source.display_name) %>%
  count(Domain, Field, Subfield, name = "n_unique_journals") # took out Subfield,

# Build ids/labels/parents for plotly sunburst
domains  <- sunburst_data %>%
  group_by(Domain) %>%
  summarise(n = sum(n_unique_journals)) %>%
  transmute(ids = Domain, labels = Domain, 
            parents = "", values = n)

fields   <- sunburst_data %>%
  group_by(Domain, Field) %>%
  summarise(n = sum(n_unique_journals), .groups = "drop") %>%
  transmute(ids = paste(Domain, Field, sep = " - "),
            labels = Field, parents = Domain, values = n)

subfields <- sunburst_data %>%
  transmute(ids     = paste(Domain, Field, Subfield, sep = " - "),
            labels  = Subfield,
            parents = paste(Domain, Field, sep = " - "),
            values  = n_unique_journals)

sb <- bind_rows(domains, fields, subfields) # , subfields

plot_ly(
  sb,
  ids     = ~ids,
  labels  = ~labels,
  parents = ~parents,
  values  = ~values,
  type    = "sunburst",
  branchvalues = "remainder"
  # branchvalues = "total"
)
```


working on a table with more fine grained analysis
Using Claude.ai Sonnet 5 Medium on August 18 to make the tree visual but will not be using it for the final paper
```{r}
# install.packages(c("igraph", "ggraph"))
library(igraph)
library(ggraph)
library(dplyr)

# 1. Start from the flattened summary (lowercase columns, per earlier fix)
# sunburst_table <- alex_doi_new1 %>%
#   filter(!is.na(doi)) %>%
#   group_by(Domain, Field, Subfield) %>%
#   summarise(doi_count = n_distinct(doi), .groups = "drop") %>%
#   arrange(Domain, Field, Subfield)

sunburst_table <- alex_doi_new1 %>%
  filter(!is.na(doi)) %>%
  mutate(
    Domain   = coalesce(Domain,   "Unclassified"),
    Field    = coalesce(Field,    "Unclassified"),
    Subfield = coalesce(Subfield, "Unclassified")
  ) %>%
  group_by(Domain, Field, Subfield) %>%
  summarise(doi_count = n_distinct(doi), .groups = "drop") %>%
  arrange(Domain, Field, Subfield)


# 2. Build unique IDs at each level so identically-named
#    fields/subfields under different domains don't collide
nodes_domain <- sunburst_table %>%
  distinct(Domain) %>%
  transmute(id = Domain, label = Domain, level = "Domain")

nodes_field <- sunburst_table %>%
  distinct(Domain, Field) %>%
  transmute(id = paste(Domain, Field, sep = " || "),
            label = Field, level = "Field")

nodes_subfield <- sunburst_table %>%
  distinct(Domain, Field, Subfield) %>%
  left_join(
    sunburst_table %>% group_by(Domain, Field, Subfield) %>%
      summarise(doi_count = sum(doi_count), .groups = "drop"),
    by = c("Domain", "Field", "Subfield")
  ) %>%
  transmute(id = paste(Domain, Field, Subfield, sep = " || "),
            label = Subfield, level = "Subfield", doi_count)

# 3. Edge list: Root -> Domain -> Field -> Subfield
edges <- bind_rows(
  sunburst_table %>% distinct(Domain) %>%
    transmute(from = "All Publications", to = Domain),
  sunburst_table %>% distinct(Domain, Field) %>%
    transmute(from = Domain, to = paste(Domain, Field, sep = " || ")),
  sunburst_table %>% distinct(Domain, Field, Subfield) %>%
    transmute(from = paste(Domain, Field, sep = " || "),
              to   = paste(Domain, Field, Subfield, sep = " || "))
)

# 4. Node table (root + all levels), with doi_count for sizing leaves
nodes <- bind_rows(
  tibble(id = "All Publications", label = "All Publications",
         level = "Root", doi_count = NA_real_),
  nodes_domain    %>% mutate(doi_count = NA_real_),
  nodes_field     %>% mutate(doi_count = NA_real_),
  nodes_subfield
) %>% distinct(id, .keep_all = TRUE)

# 5. Build graph and plot as a tidy (Reingold-Tilford) tree
g <- graph_from_data_frame(edges, vertices = nodes, directed = TRUE)

ggraph(g, layout = "tree") +
  geom_edge_diagonal(color = "grey70", width = 0.4) +
  geom_node_point(aes(size = doi_count, color = level),
                   na.rm = TRUE) +
  geom_node_text(aes(label = label, filter = level != "Subfield"),
                  size = 3, fontface = "bold", repel = TRUE) +
  geom_node_text(aes(label = label, filter = level == "Subfield"),
                  size = 2.3, angle = 60, hjust = 0, repel = TRUE) +
    # geom_node_text(aes(label = label, filter = level == "Subfield"),
    #              size = 2.3, angle = 60, hjust = 0,
    #              repel = TRUE, max.overlaps = Inf) +
  scale_color_manual(values = c(Root = "grey30", Domain = "#2c7fb8",
                                 Field = "#41b6c4", Subfield = "#a1dab4")) +
  scale_size_continuous(range = c(1, 8), guide = "none") +
  coord_flip() +
  scale_y_reverse() +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(color = "Level")
```

Tidy table
```{r}
library(dplyr)
library(tidyr)

# Helper: build "name (count); name (count); ..." for any parent -> child pairing
list_count_by <- function(data, parent_col, child_col) {
  data %>%
    filter(!is.na(doi), !is.na({{ parent_col }}), !is.na({{ child_col }})) %>%
    group_by({{ parent_col }}, {{ child_col }}) %>%
    summarise(doi_count = n_distinct(doi), .groups = "drop") %>%
    arrange({{ parent_col }}, desc(doi_count)) %>%
    group_by({{ parent_col }}) %>%
    summarise(
      n_items = n(),
      items   = paste0({{ child_col }}, " (", doi_count, ")", collapse = "; "),
      .groups = "drop"
    )
}

# Domain -> Field summary
field_summary <- alex_doi_new1 %>%
  list_count_by(Domain, Field) %>%
  rename(n_fields = n_items, fields_list = items)

# Domain -> Subfield summary (flattened straight from domain, skipping field)
subfield_summary <- alex_doi_new1 %>%
  list_count_by(Domain, Subfield) %>%
  rename(n_subfields = n_items, subfields_list = items)

# Combine into one table: Domain | # Fields | Fields (list+count) | # Subfields | Subfields (list+count)
domain_overview <- alex_doi_new1 %>%
  distinct(Domain) %>%
  filter(!is.na(Domain)) %>%
  left_join(field_summary,    by = "Domain") %>%
  left_join(subfield_summary, by = "Domain") %>%
  arrange(Domain)

print(domain_overview, width = Inf)
```
more easily rendered table
```{r}
 datatable(domain_overview, options = list(
    columnDefs = list(list(targets = c(2, 4), render = JS(
      "function(data, type, row) { return type === 'display' && data.length > 60 ? data.substr(0, 57) + '...' : data; }"
    )))
  ))


```

Tree Map
I just asked treemap in R to Google and google ai generated some generic options using cars data 8/18 
```{r}
# Plot the treemap
library(treemap)

treemap(sunburst_table,
        index = c("Domain", "Field", "Subfield"), # Hierarchical levels
        vSize = "doi_count",                        # Rectangle sizes
        type = "index",                         # How rectangles are colored
        palette = "Set2",                       # Color palette
        title = "Distribution of Publications"         # Chart title
)
```


Reading google ai mode on generic query on August 18, 2026
```{r}
# 1. Install and load packages
# install.packages(c("ggplot2", "treemapify"))
library(ggplot2)
library(treemapify)

# 2. Use the same data frame from Method 1
# 3. Build the ggplot treemap
ggplot(sunburst_table, aes(area = doi_count, fill = Domain, label = Subfield, subgroup = Domain)) +
  geom_treemap() +
  # Add white text labels for the sub-categories
  geom_treemap_text(colour = "white", place = "center", reflow = TRUE) +
  # Draw a thick border around the main categories
  geom_treemap_subgroup_border(colour = "black", size = 0.5) +
  labs(title = "Distribution of Publications by Domain and Subfields")
```



by domain, field
```{r}
sunburst_table2 <- alex_doi_new1 %>%
  filter(!is.na(doi)) %>%
  mutate(
    Domain   = coalesce(Domain,   "Unclassified"),
    Field    = coalesce(Field,    "Unclassified"),
    # Subfield = coalesce(Subfield, "Unclassified")
  ) %>%
  group_by(Domain, Field) %>%
  summarise(doi_count = n_distinct(doi), .groups = "drop") %>%
  arrange(Domain, Field)

# 2. Use the same data frame from Method 1
# 3. Build the ggplot treemap
ggplot(sunburst_table2, aes(area = doi_count, fill = Domain, label = Field, subgroup = Domain)) +
  geom_treemap() +
  # scale_fill_brewer(palette = 'Set3') +
  # Add white text labels for the sub-categories
  geom_treemap_text(colour = "black", place = "center", reflow = TRUE) +
  # Draw a thick border around the main categories
  geom_treemap_subgroup_border(colour = "darkgrey", size = 2) +
  labs(title = "Distribution of Publications by Domain and Field")

# read https://stackoverflow.com/questions/76186672/how-to-change-colors-on-treemap
```
countries represented
```{r}
# alex_doi_new_wide
DATE_MIN   <- as.Date("2016-03-15")
DATE_MAX   <- as.Date("2026-03-16")
EXCL_DOI   <- "10.1038/sdata.2016.18"

code_col <- intersect(c("code", "iso2", "iso2c", "country_code"), names(country_dict))
name_col <- intersect(c("name", "country_name", "country", "label"),  names(country_dict))

if (length(code_col) == 0) stop("country_dict must have a code column.")
country_dict_std <- if (length(name_col) == 0) {
  country_dict %>% rename(code = !!code_col[1]) %>% mutate(name = code)
} else {
  country_dict %>% rename(code = !!code_col[1], name = !!name_col[1])
}

country_codes <- names(alex_doi_new_wide)[str_detect(names(alex_doi_new_wide), "^[A-Z]{2}$")]

df_long <- alex_doi_new_wide %>%
  filter(publication_date >= DATE_MIN, publication_date <= DATE_MAX,
         doi_clean != EXCL_DOI) %>%
  mutate(publication_year = as.integer(publication_year)) %>%
  filter(!is.na(publication_year), publication_year > 2015) %>%
  select(publication_year, doi_clean, all_of(country_codes)) %>%
  pivot_longer(cols = all_of(country_codes),
               names_to = "country_code", values_to = "value") %>%
  filter(!is.na(value), value > 0) %>%
  left_join(country_dict_std, by = c("country_code" = "code")) %>%
  mutate(country_name = coalesce(name, country_code)) %>%
  select(publication_year, doi_clean, country_code, country_name, value)



```



```{r}
summary_full <- df_long %>%
  group_by(publication_year, country_code, country_name) %>%
  summarise(count = n_distinct(doi_clean), .groups = "drop") %>%
  group_by(publication_year) %>%
  mutate(year_total = sum(count)) %>%
  ungroup()

all_countries <- sort(unique(summary_full$country_name))

```

simple addition
```{r}
countries_doi_count <- df_long  %>%
  group_by(country_name) %>%
  summarise(unique_items = n_distinct(doi_clean), .groups = 'drop')


```



```{r}
df_long |>
  summarise(
    rows         = n(),
    unique_country = n_distinct(country_name),
    unique_doi = n_distinct(doi_clean),
    # Year      = publication_year,
    .groups      = "drop"
  )
```

```{r}
# subset if unique items is greater than 10
more_common_countries <- countries_doi_count %>% 
  filter(unique_items >= 10)

```

analysis using only the top 100 countries
```{r}
more_common_countries_list <- more_common_countries$country_name

test_countries <- df_long %>% filter(country_name %in% more_common_countries_list)


test_countries |>
  summarise(
    rows         = n(),
    unique_country = n_distinct(country_name),
    unique_doi = n_distinct(doi_clean),
    # Year      = publication_year,
    .groups      = "drop"
  )
```

```{r}
library(gtsummary)
library(gt)

most_active_countries <- countries_doi_count %>% 
  filter(unique_items >= 1000)
# Create a publication-ready table 1

most_active_countries_table <- most_active_countries|> 
  arrange(desc(unique_items)) |> 
  gt() |> 
  cols_label(
    country_name = "Country of Author(s)",
    unique_items  = "Unique DOI's",
    # wt  = "Weight (1000 lbs)"
  ) |> 
  tab_header(
    title = "Top 10 Most Active Countries of Authorship",
    subtitle = "Base on OpenAlex"
  ) 

# my_table <- mtcars |> 
#   select(mpg, hp, wt) |> 
#   head() |> 
#   arrange(hp) |>  # Sorter by hp (lowest to highest)
#   gt() |> 
#   cols_label(
#     mpg = "Miles Per Gallon",
#     hp  = "Horsepower",
#     wt  = "Weight (1000 lbs)"
#   )

most_active_countries_table
```

```{r}
gtsave(most_active_countries_table, filename = "visual_output/most_active_countries_table.docx")

```


```{r}
most_active_countries_list <- most_active_countries$country_name

test1_countries <- df_long %>% filter(country_name %in% most_active_countries_list)


test1_countries |>
  summarise(
    rows         = n(),
    unique_country = n_distinct(country_name),
    unique_doi = n_distinct(doi_clean),
    # Year      = publication_year,
    .groups      = "drop"
  )
```

### journal hierarchy ###
downloaded from https://www.scimagojr.com/journalrank.php
first downloaded april 23, 2026
```{r}
#### SECTION with JOURNAL PRESTIGE ####
# read in Scimagojr data
library(readr)
scimagojr <- read_delim("scimagojr_data/scimagojr 2025.csv",
                             delim = ";", escape_double = FALSE, trim_ws = TRUE)
scimagojr <- read_delim(
  "scimagojr_data/scimagojr 2025.csv",
  delim          = ";",
  locale         = locale(decimal_mark = ","),
  show_col_types = FALSE
)

# View(scimagojr)

# subset scimagojr
scimagojr_trim <- scimagojr %>% 
  select(Title, Region, Categories, Areas, Overton, Issn,
         `H index`, `SJR Best Quartile`, `Citations / Doc. (2years)`,
         `Total Refs.`, `Total Citations (3years)`, `Total Docs. (2025)`, 
         `Total Docs. (3years)`, Rank
  )



```


```{r}
library(stringr)

normalize <- function(x) {
  x %>%
    str_to_lower() %>%
    str_remove_all("^the\\s+") %>%
    str_remove_all("[^a-z0-9 ]") %>%
    str_squish()
}

alex_clean    <- alex_doi_new %>% mutate(title_key = normalize(primary_location.source.display_name))
scimago_clean <- scimagojr_trim    %>% mutate(title_key = normalize(Title))

alex_doi_new_journal <- alex_clean %>%
  left_join(scimago_clean, by = "title_key")

# --- Pivot ISSN column into long format (one ISSN per row) ---
scimagojr_issn <- scimagojr %>%
  mutate(Issn = str_remove_all(Issn, "\\s")) %>%    # strip whitespace
  separate_rows(Issn, sep = ",") %>%                 # one row per ISSN
  filter(Issn != "", !is.na(Issn)) %>%               # drop blanks
  mutate(Issn = str_squish(Issn))

# --- Join on ISSN ---
# OpenAlex typically stores ISSNs as "XXXX-XXXX" so normalise both sides
normalize_issn <- function(x) str_remove_all(str_to_upper(x), "-")

alex_doi_new_journal <- alex_doi_new %>%
  mutate(issn_key = normalize_issn(primary_location.source.issn_l)) %>%
  left_join(
    scimagojr_issn %>% mutate(issn_key = normalize_issn(Issn)),
    by = "issn_key"
  )

# write csv
# write.csv(alex_doi_new_journal, file = "output_data/alex_doi_new_journal.csv")

# # table with all the journals and counts 
# n_journal <- n_distinct(alex_doi_new_journal$primary_location.source.display_name)

journal_counts <- alex_doi_new_journal %>%
  distinct(doi_clean, primary_location.source.display_name) %>% 
  count(primary_location.source.display_name, name = "Count") %>%
  select(primary_location.source.display_name, Count) %>% 
  rename("Journal Title" = primary_location.source.display_name) %>% 
  arrange(desc(Count))

journal_counts


```


```{r}
# interactive graph
library(dplyr)
library(tidyr)

journal_year_counts <- alex_doi_new_journal %>%
  filter(!is.na(primary_location.source.display_name),
         !is.na(publication_year)) %>%
  distinct(
    doi_clean,
    primary_location.source.display_name,
    publication_year,
    `SJR Best Quartile`
    # `H index`
    # domain
  ) %>%
  count(
    primary_location.source.display_name,
    publication_year,
    `SJR Best Quartile`,
    # `H index`,
    # domain,
    name = "Citations"
  ) %>%
  rename(
    `Journal Title` = primary_location.source.display_name,
    Year = publication_year
  ) %>%
  arrange(`Journal Title`, Year)
```


```{r}
library(DT)

datatable(
  journal_year_counts,
  extensions = c("Buttons"),
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    order = list(list(0, "asc"), list(1, "asc"))
  ),
  filter = "top",
  rownames = FALSE
)

# # # left join title to journal title
# alex_doi_new_journal <- alex_doi_new %>%
#   left_join(scimagojr, by = c("primary_location.source.display_name" = "Title"))

# See how many failed to match
alex_doi_new_journal %>% 
  filter(is.na(`SJR Best Quartile`)) %>% 
  distinct(primary_location.source.display_name) %>% 
  nrow()

# unmatched journals
unmatched_journals <- alex_doi_new_journal %>% 
  filter(is.na(`SJR Best Quartile`)) %>% 
  distinct(primary_location.source.display_name) 
```

data on SJR best quartiles
```{r}

```

