---
title: "Open_alex_static_visuals"
format: html
editor: visual
---

Sunburst chart August 17, 2026 modified after using a new dataset that merged from open alex, downloaded August 7, 2026

create tables with this information
```{r}
library(readr)
library(tidyverse)
alex_doi_new1 <- read_csv("output_data/alex_doi_new1.csv")
```



```{r}
library(DT)

journal_hierarchy <- alex_doi_new1 %>%
  filter(!is.na(Domain), 
         # !is.na(field), !is.na(subfield),
         !is.na(publication_year),
         !is.na(primary_location.source.display_name)) %>%
  distinct(publication_year, Domain, Field, Subfield,
           primary_location.source.display_name) %>%
  count(publication_year, Domain, Field, Subfield,
        name = "n_unique_journals") %>%
  arrange(publication_year, Domain, Field, Subfield)

datatable(
  journal_hierarchy,
  filter   = "top",           # per-column filter boxes
  rownames = FALSE,
  colnames = c("Year", "Domain", "Field", "Subfield", "Unique Journals"),
  options  = list(
    pageLength  = 20,
    scrollX     = TRUE,
    autoWidth   = TRUE,
    columnDefs  = list(list(className = "dt-center", targets = c(0, 4)))
  )
)

```
clean up the names of domain, field, subfield
```{r}
# Summary table grouped by the unique DOI
domain_summary <- alex_doi_new1 %>%
  filter(Domain !="") %>% 
  group_by(Domain) %>%
  summarise(
    unique_dois_count = n_distinct(doi, na.rm = TRUE),
    unique_field_count = n_distinct(Field, na.rm = TRUE),
    unique_subfield_count = n_distinct(Subfield, na.rm = TRUE)
  )


# Print the resulting table
print(domain_summary)
```

```{r}
# Create a flattened hierarchical table
sunburst_table <- alex_doi_new1 %>%
  group_by(Domain, Field, Subfield) %>%
  summarise(
    doi_count = n_distinct(doi, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Sort from top tier down to make the hierarchy easy to read
  arrange(Domain, Field, Subfield)

print(sunburst_table)
```

```{r}
# Create a clean summary table
domain_counts <- alex_doi_new1 %>%
  group_by(Domain) %>%
  summarise(
    doi_count = n_distinct(doi, na.rm = TRUE),
    .groups = "drop",
    field_count = n_distinct(Field),  # Counts unique schools
    subfield_count = n()                                   # Counts total sports rows
  )

grid_counts <- alex_doi_new1 %>%
  count(Field, Subfield) %>%
  pivot_wider(names_from = Subfield, values_from = n, values_fill = 0)

print(grid_counts)
```

# attempt a dendogram
```{r}
# Libraries
library(ggraph)
library(igraph)
library(tidyverse)
theme_set(theme_void())


```


```{r}
# Plotly Sunburst visual
library(plotly)
library(dplyr)

sunburst_data <- alex_doi_new1 %>%
  filter(!is.na(Domain), 
         # !is.na(field), !is.na(subfield),
         !is.na(primary_location.source.display_name)) %>%
  distinct(Domain, Field, 
           # Subfield,
           primary_location.source.display_name) %>%
  count(Domain, Field, name = "n_unique_journals") # took out Subfield,

# Build ids/labels/parents for plotly sunburst
domains  <- sunburst_data %>%
  group_by(Domain) %>%
  summarise(n = sum(n_unique_journals)) %>%
  transmute(ids = Domain, labels = Domain, 
            parents = "", values = n)

fields   <- sunburst_data %>%
  group_by(Domain, Field) %>%
  summarise(n = sum(n_unique_journals), .groups = "drop") %>%
  transmute(ids = paste(Domain, Field, sep = " - "),
            labels = Field, parents = Domain, values = n)

# subfields <- sunburst_data %>%
#   transmute(ids     = paste(Domain, Field, Subfield, sep = " - "),
#             labels  = Subfield,
#             parents = paste(Domain, Field, sep = " - "),
#             values  = n_unique_journals)

sb <- bind_rows(domains, fields) # , subfields

plot_ly(
  sb,
  ids     = ~ids,
  labels  = ~labels,
  parents = ~parents,
  values  = ~values,
  type    = "sunburst",
  branchvalues = "remainder"
  # branchvalues = "total"
)
```

```{r}
table(alex_doi_new1$Domain)
```

make a publishable table asked google ai with generic code august 20, 2026
```{r}
library(gtsummary)
library(gt)

# Generate the table including unclassified (NA) values
# (Replacing your original command with an explicit useNA argument)
domain_table <- table(alex_doi_new1$Domain, useNA = "ifany")

# 2. Convert the table into a clean data frame for formatting
table_df <- as.data.frame(domain_table)
colnames(table_df) <- c("Domain", "Frequency")

# 3. Clean up the "unclassified" label if it appears as <NA>
table_df$Domain <- as.character(table_df$Domain)
table_df$Domain[is.na(table_df$Domain)] <- "Unclassified"

domain_freq_table <- table_df|> 
  arrange(desc(Frequency)) |> 
  gt() |> 
  cols_align(align = "left", columns = Domain) %>%
  cols_align(align = "right", columns = Frequency) %>%
    # wt  = "Weight (1000 lbs)"
  tab_header(
    title = "Research Domains and the Unique DOI's",
    subtitle = "Base on OpenAlex"
  ) 

# 4. Build and style the professional table
domain_freq_table <- table_df %>%
  gt() %>%
  # Align text left and numbers right
  cols_align(align = "left", columns = Domain) %>%
  cols_align(align = "right", columns = Frequency) %>%
  cols_label(
    Domain= "Domain",
    Frequency  = "Unique DOI's",
  ) 
  # # Bold the column headers
  # tab_style(
  #   style = cell_text(weight = "bold"),
  #   locations = cells_column_labels()
  # )

# Print the table
domain_freq_table
# countries_doi_count <- df_long  %>%
#   group_by(country_name) %>%
#   summarise(unique_items = n_distinct(doi_clean), .groups = 'drop')
# 
# most_active_countries <- countries_doi_count %>% 
#   filter(unique_items >= 1000)
# # Create a publication-ready table 1
# 
# most_active_countries_table <- most_active_countries|> 
#   arrange(desc(unique_items)) |> 
#   gt() |> 
#   cols_label(
#     country_name = "Country of Author(s)",
#     unique_items  = "Unique DOI's",
#     # wt  = "Weight (1000 lbs)"
#   ) |> 
#   tab_header(
#     title = "Top 10 Most Active Countries of Authorship",
#     subtitle = "Base on OpenAlex"
#   ) 
gtsave(domain_freq_table, filename = "visual_output/domain_freq_table.docx")
```



```{r}
sunburst_data <- alex_doi_new1 %>%
  filter(!is.na(Domain), 
         # !is.na(field), !is.na(subfield),
         !is.na(primary_location.source.display_name)) %>%
  distinct(Domain, Field, Subfield,
           primary_location.source.display_name) %>%
  count(Domain, Field, Subfield, name = "n_unique_journals") # took out Subfield,

# Build ids/labels/parents for plotly sunburst
domains  <- sunburst_data %>%
  group_by(Domain) %>%
  summarise(n = sum(n_unique_journals)) %>%
  transmute(ids = Domain, labels = Domain, 
            parents = "", values = n)

fields   <- sunburst_data %>%
  group_by(Domain, Field) %>%
  summarise(n = sum(n_unique_journals), .groups = "drop") %>%
  transmute(ids = paste(Domain, Field, sep = " - "),
            labels = Field, parents = Domain, values = n)

subfields <- sunburst_data %>%
  transmute(ids     = paste(Domain, Field, Subfield, sep = " - "),
            labels  = Subfield,
            parents = paste(Domain, Field, sep = " - "),
            values  = n_unique_journals)

sb <- bind_rows(domains, fields, subfields) # , subfields

plot_ly(
  sb,
  ids     = ~ids,
  labels  = ~labels,
  parents = ~parents,
  values  = ~values,
  type    = "sunburst",
  branchvalues = "remainder"
  # branchvalues = "total"
)
```


working on a table with more fine grained analysis
Using Claude.ai Sonnet 5 Medium on August 18 to make the tree visual but will not be using it for the final paper
```{r}
# install.packages(c("igraph", "ggraph"))
library(igraph)
library(ggraph)
library(dplyr)

# 1. Start from the flattened summary (lowercase columns, per earlier fix)
# sunburst_table <- alex_doi_new1 %>%
#   filter(!is.na(doi)) %>%
#   group_by(Domain, Field, Subfield) %>%
#   summarise(doi_count = n_distinct(doi), .groups = "drop") %>%
#   arrange(Domain, Field, Subfield)

sunburst_table <- alex_doi_new1 %>%
  filter(!is.na(doi)) %>%
  mutate(
    Domain   = coalesce(Domain,   "Unclassified"),
    Field    = coalesce(Field,    "Unclassified"),
    Subfield = coalesce(Subfield, "Unclassified")
  ) %>%
  group_by(Domain, Field, Subfield) %>%
  summarise(doi_count = n_distinct(doi), .groups = "drop") %>%
  arrange(Domain, Field, Subfield)


# 2. Build unique IDs at each level so identically-named
#    fields/subfields under different domains don't collide
nodes_domain <- sunburst_table %>%
  distinct(Domain) %>%
  transmute(id = Domain, label = Domain, level = "Domain")

nodes_field <- sunburst_table %>%
  distinct(Domain, Field) %>%
  transmute(id = paste(Domain, Field, sep = " || "),
            label = Field, level = "Field")

nodes_subfield <- sunburst_table %>%
  distinct(Domain, Field, Subfield) %>%
  left_join(
    sunburst_table %>% group_by(Domain, Field, Subfield) %>%
      summarise(doi_count = sum(doi_count), .groups = "drop"),
    by = c("Domain", "Field", "Subfield")
  ) %>%
  transmute(id = paste(Domain, Field, Subfield, sep = " || "),
            label = Subfield, level = "Subfield", doi_count)

# 3. Edge list: Root -> Domain -> Field -> Subfield
edges <- bind_rows(
  sunburst_table %>% distinct(Domain) %>%
    transmute(from = "All Publications", to = Domain),
  sunburst_table %>% distinct(Domain, Field) %>%
    transmute(from = Domain, to = paste(Domain, Field, sep = " || ")),
  sunburst_table %>% distinct(Domain, Field, Subfield) %>%
    transmute(from = paste(Domain, Field, sep = " || "),
              to   = paste(Domain, Field, Subfield, sep = " || "))
)

# 4. Node table (root + all levels), with doi_count for sizing leaves
nodes <- bind_rows(
  tibble(id = "All Publications", label = "All Publications",
         level = "Root", doi_count = NA_real_),
  nodes_domain    %>% mutate(doi_count = NA_real_),
  nodes_field     %>% mutate(doi_count = NA_real_),
  nodes_subfield
) %>% distinct(id, .keep_all = TRUE)

# 5. Build graph and plot as a tidy (Reingold-Tilford) tree
g <- graph_from_data_frame(edges, vertices = nodes, directed = TRUE)

ggraph(g, layout = "tree") +
  geom_edge_diagonal(color = "grey70", width = 0.4) +
  geom_node_point(aes(size = doi_count, color = level),
                   na.rm = TRUE) +
  geom_node_text(aes(label = label, filter = level != "Subfield"),
                  size = 3, fontface = "bold", repel = TRUE) +
  geom_node_text(aes(label = label, filter = level == "Subfield"),
                  size = 2.3, angle = 60, hjust = 0, repel = TRUE) +
    # geom_node_text(aes(label = label, filter = level == "Subfield"),
    #              size = 2.3, angle = 60, hjust = 0,
    #              repel = TRUE, max.overlaps = Inf) +
  scale_color_manual(values = c(Root = "grey30", Domain = "#2c7fb8",
                                 Field = "#41b6c4", Subfield = "#a1dab4")) +
  scale_size_continuous(range = c(1, 8), guide = "none") +
  coord_flip() +
  scale_y_reverse() +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(color = "Level")
```

Tidy table
```{r}
library(dplyr)
library(tidyr)

# Helper: build "name (count); name (count); ..." for any parent -> child pairing
list_count_by <- function(data, parent_col, child_col) {
  data %>%
    filter(!is.na(doi), !is.na({{ parent_col }}), !is.na({{ child_col }})) %>%
    group_by({{ parent_col }}, {{ child_col }}) %>%
    summarise(doi_count = n_distinct(doi), .groups = "drop") %>%
    arrange({{ parent_col }}, desc(doi_count)) %>%
    group_by({{ parent_col }}) %>%
    summarise(
      n_items = n(),
      items   = paste0({{ child_col }}, " (", doi_count, ")", collapse = "; "),
      .groups = "drop"
    )
}

# Domain -> Field summary
field_summary <- alex_doi_new1 %>%
  list_count_by(Domain, Field) %>%
  rename(n_fields = n_items, fields_list = items)

# Domain -> Subfield summary (flattened straight from domain, skipping field)
subfield_summary <- alex_doi_new1 %>%
  list_count_by(Domain, Subfield) %>%
  rename(n_subfields = n_items, subfields_list = items)

# Combine into one table: Domain | # Fields | Fields (list+count) | # Subfields | Subfields (list+count)
domain_overview <- alex_doi_new1 %>%
  distinct(Domain) %>%
  filter(!is.na(Domain)) %>%
  left_join(field_summary,    by = "Domain") %>%
  left_join(subfield_summary, by = "Domain") %>%
  arrange(Domain)

print(domain_overview, width = Inf)
```
more easily rendered table
```{r}
 datatable(domain_overview, options = list(
    columnDefs = list(list(targets = c(2, 4), render = JS(
      "function(data, type, row) { return type === 'display' && data.length > 60 ? data.substr(0, 57) + '...' : data; }"
    )))
  ))


```

Tree Map
I just asked treemap in R to Google and google ai generated some generic options using cars data 8/18 
```{r}
# Plot the treemap
library(treemap)

treemap(sunburst_table,
        index = c("Domain", "Field", "Subfield"), # Hierarchical levels
        vSize = "doi_count",                        # Rectangle sizes
        type = "index",                         # How rectangles are colored
        palette = "Set2",                       # Color palette
        title = "Distribution of Publications"         # Chart title
)
```


Reading google ai mode on generic query on August 18, 2026 query for treemap in R.
code updated with a prompt to claude ai sonnet 5 for prompts on web accessible colors and min.size on August 31, 2026.
```{r}
# 1. Install and load packages
# install.packages(c("showtext", "sysfonts"))
library(ggplot2)
library(treemapify)
library(extrafont)
# font_import()   # one-time, slow — scans system fonts
# loadfonts(device = "win")
library(showtext) 
#face="bold",
library(sysfonts) 
#for family = "mono" for text

okabe_ito <- c("#E69F00", "#56B4E9", "#009E73","#F0E442", "#0072B2")  # 4 colors, use as many as you need
# 2. Use the same data frame from Method 1
# 3. Build the ggplot treemap
ggplot(sunburst_table, aes(area = doi_count, fill = Domain, label = Subfield, subgroup = Domain)) +
  geom_treemap() +
  # Add white text labels for the sub-categories
  geom_treemap_text(colour = "black", place = "center", reflow = TRUE,size = 14, 
                    # fontface = "italic", 
                    family = "serif",
                    min.size = 6) +
  # Draw a thick border around the main categories
  geom_treemap_subgroup_border(colour = "darkgrey", size = 0.5) +
    scale_fill_manual(values = okabe_ito) +
  labs(title = "Distribution of Publications by Domain and Subfields") +
    theme(
    plot.title = element_text(family = "serif", size = 16, face = "bold"),
    text = element_text(family = "serif")   
  )


```

adding a helping column per conversation with claude 8/31/26
```{r}
library(dplyr)

# Flag the top subfield within each domain, and build a label
# that shows "Subfield (n = count)" only for that one
sunburst_table <- sunburst_table %>%
  group_by(Domain) %>%
  mutate(
    is_top_subfield = doi_count == max(doi_count),
    # shortened display name, only for labeling — keeps Subfield itself untouched
subfield_short = case_when(
  Subfield == "Public Health, Environmental and Occupational Health" ~ "Public, Environmental, and Occupational Health",
  Subfield == "Management Science and Operations Research"    ~ "Management Sci. and Operations Research",
  Subfield == "Economics, Econometrics and Finance"           ~ "Econ/Econometrics/Finance",
  TRUE ~ Subfield
),
    label_text = if_else(
      is_top_subfield,
      paste0(subfield_short, "\n(n = ", scales::comma(doi_count), ")"),
      subfield_short
    )
  ) %>%
  ungroup()
# sunburst_table <- sunburst_table %>%
#   group_by(Domain) %>%
#   mutate(
#     is_top_subfield = doi_count == max(doi_count),
#     label_text = if_else(
#       is_top_subfield,
#       paste0(Subfield, "\n(n = ", scales::comma(doi_count), ")"),
#       Subfield
#     )
#   ) %>%
#   ungroup()

okabe_ito <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")

ggplot(sunburst_table, aes(area = doi_count, fill = Domain, label = label_text, subgroup = Domain)) +
  geom_treemap() +
  geom_treemap_text(colour = "black", place = "center", reflow = TRUE,
                     # family = "serif",
                     min.size = 5) +
  geom_treemap_subgroup_border(colour = "darkgrey", size = 0.5) +
  scale_fill_manual(values = okabe_ito) +
  labs(title = "Distribution of Publications by Domain and Subfields") 
  # theme(
  #   plot.title = element_text(family = "serif", size = 12, face = "bold"),
  #   text = element_text(family = "serif")
  # )
```
```{r}
sunburst_table %>%
  group_by(Domain) %>%
  mutate(pct_of_domain = doi_count / sum(doi_count) * 100,
         pct_of_total  = doi_count / sum(sunburst_table$doi_count) * 100) %>%
  filter(is_top_subfield) %>%
  select(Domain, Subfield, doi_count, pct_of_domain, pct_of_total)
```


# saturation of the papers by open alex existing domains
All four domains
## within each domain, count the fields
```{r}
library(usethis)
library(openalexR)
# View the data structure
# dplyr::glimpse(paper_data)

###### do not share #################################
usethis::edit_r_environ()

options(openalexR.apikey = "")

# Sys.getenv("OPENALEX_API_KEY")
# # fetch paper info
paper_data <- oa_fetch(
  identifier = "W2302501749", # OpenAlex ID for the orignal paper is W2302501749
  entity = "works",
  verbose = TRUE
)
```
Google search on api key for OpenAlex to get field data on 8/24/2026 I got back
"4. Example: openalexR (R Package)If you are developing in R, you can add your key to your .Rprofile or .Renviron file using openalexR.apikey = "YOUR_API_KEY", then restrict your data frames using the select configuration:rlibrary(openalexR)

oa_fetch(
  entity = "works",
  search = "machine learning",
  options = oa_options(select = c("id", "display_name", "authorships"))
)
Use code with caution.If you need a breakdown of what fields are available for a particular endpoint, let me know which entity type (e.g., Works, Authors, Institutions, or Sources) you are targeting!"
additional prompt - how do I write the API code in R? In the API The Domains endpoint is at api.openalex.org/domains. Fetch a single domain by ID — /domains/1 — or a list; with only four domains, one unfiltered request returns them all. You can still filter, search, sort, group, and page over them. You can filter and group domains by fields.id, and by works_count, cited_by_count, id, and display_name, all of which also sort. Full-text matching uses the search parameter (the older display_name.search filter is deprecated). To find works in a domain, use the Works endpoint: filter=primary_topic.domain.id:1 (primary only) or filter=topics.domain.id:1 (any assigned topic). For the full list of endpoints see the endpoints index.

```{r}
# library(httr2)
# library(jsonlite)
# library(dplyr)
# 
# # 1. Provide your working OpenAlex API Key
# my_api_key <- "MMM"
# 
# # 2. Build a universally compliant URL string for the flat fields endpoint
# # Since there are only 26 fields, a per_page of 50 captures all data in one response
# target_url <- paste0("https://openalex.org", my_api_key, "&per_page=50")
# 
# # 3. Read and parse the JSON directly from the API endpoint
# fields_data <- jsonlite::fromJSON(target_url)
# fields_df <- fields_data$results
# 
# # 4. Extract and map the 26 fields cleanly to their parent domains
# mapped_fields <- data.frame(
#   domain_name = fields_df$domain$display_name,
#   domain_id   = fields_df$domain$id,
#   field_name  = fields_df$display_name,
#   field_id    = fields_df$id
# ) |> 
#   dplyr::arrange(domain_name, field_name)
# 
# # 5. Loop and print out the complete structured table 
# print(mapped_fields, max = Inf)
# 
# # 6. Sanity check validation counter
# cat("\nTotal rows successfully processed:", nrow(mapped_fields), "\n")
```
failed to work

documentation from OpenAlex as of August 11, 2026

Signal	# groups	Familiarity	Fit to custom areas
SDGs	17	High	Low
Domains	4	High	Low
Fields	26	High	Low
Subfields	252	High	Medium
Topics	4,516	Low	Medium-high

```{r}
summary(alex_doi_new1)
```
We have coverage in all domains and fields. and 210 out of 252 subfields in OpenAlex.
Find which ones are the missing subfields 
claude.ai prompt for Sonnet 5 Medium on Aug. 24, 2026 - I asked create a API pull in OpenAlex using R to get subfields data. there are 252 subfields I want to know what their fields are. In the API
The Subfields endpoint is at `api.openalex.org/subfields`. Fetch a single subfield by ID — `/subfields/2718` — or a list, and [filter](https://help.openalex.org/api/filtering/), [search](https://help.openalex.org/api/searching/), [sort](https://help.openalex.org/api/sorting/), [group](https://help.openalex.org/api/grouping/), and [page](https://help.openalex.org/api/paging/) over it.
You can filter and group subfields by their place in the hierarchy — `domain.id`, `field.id`, and `topics.id` — and by `works_count`, `cited_by_count`, `id`, and `display_name`, all of which also sort. Full-text matching uses the `search`[ parameter](https://help.openalex.org/api/searching/) (the older `display_name.search` filter is deprecated). To find works in a subfield, use the [Works](https://help.openalex.org/data/works/) endpoint: `filter=primary_topic.subfield.id:2718` (primary only) or `filter=topics.subfield.id:2718` (any assigned topic). For the full list of endpoints see the [endpoints index](https://help.openalex.org/api/endpoints/).
Last updated August 11, 2026 (this was from the openalex website)
response code
```{r}
library(httr2)
library(dplyr)
library(purrr)
library(tibble)

# Pull all subfields via cursor pagination (per_page max = 200,
# and there are 252 subfields, so this will take 2 pages)
get_openalex_subfields <- function(mailto = "your_email@example.com") {
  
  base_url <- "https://api.openalex.org/subfields"
  cursor   <- "*"
  all_results <- list()
  
  repeat {
    resp <- request(base_url) %>%
      req_url_query(
        per_page = 200,
        cursor   = cursor,
        mailto   = mailto          # polite pool: faster, more reliable rate limits
      ) %>%
      req_perform()
    
    parsed <- resp_body_json(resp, simplifyVector = FALSE)
    
    all_results <- c(all_results, parsed$results)
    
    cursor <- parsed$meta$next_cursor
    if (is.null(cursor) || length(parsed$results) == 0) break
  }
  
  all_results
}

subfields_raw <- get_openalex_subfields(mailto = "your_email@example.com")
length(subfields_raw)   # should be 252

# Flatten into a tidy data frame: subfield -> its parent field -> its parent domain
subfields_df <- map_dfr(subfields_raw, function(x) {
  tibble(
    subfield_id    = x$id,
    subfield_name  = x$display_name,
    field_id       = x$field$id,
    field_name     = x$field$display_name,
    domain_id      = x$domain$id,
    domain_name    = x$domain$display_name,
    works_count    = x$works_count %||% NA_integer_,
    cited_by_count = x$cited_by_count %||% NA_integer_
  )
})

# Strip the "https://openalex.org/" prefix off the IDs so you're left with the
# bare numeric/alpha IDs, matching the style you used earlier (e.g. "2718")
subfields_df <- subfields_df %>%
  mutate(
    subfield_id = sub("https://openalex.org/subfields/", "", subfield_id),
    field_id    = sub("https://openalex.org/fields/",    "", field_id),
    domain_id   = sub("https://openalex.org/domains/",   "", domain_id)
  ) %>%
  arrange(domain_name, field_name, subfield_name)

print(subfields_df, n = 20)
```
from Claude - A few notes:

mailto parameter: swap in a real email address. It's optional but puts you in OpenAlex's "polite pool," which gets faster/more consistent response times and is the recommended practice per their docs.
per_page = 200 is the API max, and cursor-based paging (cursor = "*" then following meta$next_cursor) is required for list endpoints beyond the first page — offset-based paging caps out and isn't reliable for full-dataset pulls.
The %||% (null-coalescing) operator comes from purrr/rlang — handles any subfield missing a works_count/cited_by_count gracefully instead of erroring.
This gives you exactly the field ↔ subfield ↔ domain mapping table you asked about — subfields_df %>% count(domain_name, field_name) will show you how many subfields sit under each field.

Since this maps directly onto the domain/field/subfield hierarchy already in alex_doi_new1, do you want to:

Save this as a lookup CSV to join against your existing data (e.g. to double check your works CSV's Domain/Field/Subfield assignments are internally consistent with OpenAlex's canonical hierarchy), or
Use it to build controlled dropdown choices (field_choices, subfield_choices) in the Shiny app instead of deriving them from whatever appears in your data?


# modified code
```{r}
library(httr2)
library(dplyr)
library(purrr)
library(tibble)

# Pull all subfields via cursor pagination (per_page max = 200,
# and there are 252 subfields, so this will take 2 pages)
get_openalex_subfields <- function(mailto = "jungmee.park@wichita.edu") {
  
  base_url <- "https://api.openalex.org/subfields"
  cursor   <- "*"
  all_results <- list()
  
  repeat {
    resp <- request(base_url) %>%
      req_url_query(
        per_page = 200,
        cursor   = cursor,
        mailto   = mailto          # polite pool: faster, more reliable rate limits
      ) %>%
      req_perform()
    
    parsed <- resp_body_json(resp, simplifyVector = FALSE)
    
    all_results <- c(all_results, parsed$results)
    
    cursor <- parsed$meta$next_cursor
    if (is.null(cursor) || length(parsed$results) == 0) break
  }
  
  all_results
}

subfields_raw <- get_openalex_subfields(mailto = "jungmee.park@wichita.edu")
length(subfields_raw)   # should be 252

# Flatten into a tidy data frame: subfield -> its parent field -> its parent domain
subfields_df <- map_dfr(subfields_raw, function(x) {
  tibble(
    subfield_id    = x$id,
    subfield_name  = x$display_name,
    field_id       = x$field$id,
    field_name     = x$field$display_name,
    domain_id      = x$domain$id,
    domain_name    = x$domain$display_name,
    works_count    = x$works_count %||% NA_integer_,
    cited_by_count = x$cited_by_count %||% NA_integer_
  )
})

# Strip the "https://openalex.org/" prefix off the IDs so you're left with the
# bare numeric/alpha IDs, matching the style you used earlier (e.g. "2718")
subfields_df <- subfields_df %>%
  mutate(
    subfield_id = sub("https://openalex.org/subfields/", "", subfield_id),
    field_id    = sub("https://openalex.org/fields/",    "", field_id),
    domain_id   = sub("https://openalex.org/domains/",   "", domain_id)
  ) %>%
  arrange(domain_name, field_name, subfield_name)

print(subfields_df, n = 20)
```
Google ai prompt. dataframe A and dataframe b compare the different responses in the same columns (column names are slightly different) in R (8/24/2026)

```{r}
# response code
# # Example if columns are shuffled
# library(dplyr)
# 
# df_B <- df_B %>% 
#   rename(col1 = slight_diff_name1, col2 = slight_diff_name2) %>% 
#   select(colnames(df_A))

subfields_df <- subfields_df |> 
  mutate(Subfield = subfield_name)

library(waldo)

subfields_df_A <- subfields_df$Subfield
subfields_df_B <- alex_doi_new1$Subfield

# Elements in A that are completely missing from B
unshown_subfields <- setdiff(subfields_df_A, subfields_df_B)
```

Google ai prompt "can you compare two variables in two different datasets" on 8/24/2026
response 
<!-- # Extract the two categorical variables -->
<!-- cat_A <- df_A$var1 -->
<!-- cat_B <- df_B$var2 -->

<!-- # Calculate percentage of exact matches (ignoring NAs) -->
<!-- mean(cat_A == cat_B, na.rm = TRUE) -->

<!-- # View a breakdown of mismatches using a cross-tabulation -->
<!-- table(cat_A, cat_B, useNA = "ifany") -->

look at subfields_df for unshown fields
```{r}
unknown_fields <- subfields_df |> 
  filter(Subfield %in% unshown_subfields)


summary(unknown_fields)
library(treemapify)
library(ggplot2)

ggplot(unknown_fields, aes(area = works_count, fill = domain_name,
                            label = subfield_name, subgroup = domain_name)) +
  geom_treemap() +
  geom_treemap_subgroup_border(color = "white", size = 2) +
  geom_treemap_subgroup_text(place = "topleft", grow = FALSE,
                              alpha = 0.6, colour = "black",
                              fontface = "italic", size = 9) +
  geom_treemap_text(colour = "white", place = "center",
                     grow = FALSE, size = 8, reflow = TRUE) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Subfields by works_count, grouped by field",
       fill = "Domain") +
  theme(legend.position = "bottom")

```

```{r}
library(reactable)

reactable(
  unknown_fields %>% select(domain_name, field_name, subfield_name, works_count, cited_by_count),
  groupBy = c("domain_name", "field_name"),
  defaultExpanded = TRUE,   # fine to leave expanded now — only 35 rows total
  columns = list(
    works_count    = colDef(aggregate = "sum", format = colFormat(separators = TRUE)),
    cited_by_count = colDef(aggregate = "sum", format = colFormat(separators = TRUE))
  )
)
```



```{r}
sunburst_table2 <- alex_doi_new1 %>%
  filter(!is.na(doi)) %>%
  mutate(
    Domain   = coalesce(Domain,   "Unclassified"),
    Field    = coalesce(Field,    "Unclassified"),
    # Subfield = coalesce(Subfield, "Unclassified")
  ) %>%
  group_by(Domain, Field) %>%
  summarise(doi_count = n_distinct(doi), .groups = "drop") %>%
  arrange(Domain, Field)

# 2. Use the same data frame from Method 1
# 3. Build the ggplot treemap
ggplot(sunburst_table2, aes(area = doi_count, fill = Domain, label = Field, subgroup = Domain)) +
  geom_treemap() +
  # scale_fill_brewer(palette = 'Set3') +
  # Add white text labels for the sub-categories
  geom_treemap_text(colour = "black", place = "center", reflow = TRUE) +
  # Draw a thick border around the main categories
  geom_treemap_subgroup_border(colour = "darkgrey", size = 2) +
  labs(title = "Distribution of Publications by Domain and Field")

# read https://stackoverflow.com/questions/76186672/how-to-change-colors-on-treemap
```
countries represented
```{r}
# alex_doi_new_wide
DATE_MIN   <- as.Date("2016-03-15")
DATE_MAX   <- as.Date("2026-03-16")
EXCL_DOI   <- "10.1038/sdata.2016.18"

code_col <- intersect(c("code", "iso2", "iso2c", "country_code"), names(country_dict))
name_col <- intersect(c("name", "country_name", "country", "label"),  names(country_dict))

if (length(code_col) == 0) stop("country_dict must have a code column.")
country_dict_std <- if (length(name_col) == 0) {
  country_dict %>% rename(code = !!code_col[1]) %>% mutate(name = code)
} else {
  country_dict %>% rename(code = !!code_col[1], name = !!name_col[1])
}

country_codes <- names(alex_doi_new_wide)[str_detect(names(alex_doi_new_wide), "^[A-Z]{2}$")]

df_long <- alex_doi_new_wide %>%
  filter(publication_date >= DATE_MIN, publication_date <= DATE_MAX,
         doi_clean != EXCL_DOI) %>%
  mutate(publication_year = as.integer(publication_year)) %>%
  filter(!is.na(publication_year), publication_year > 2015) %>%
  select(publication_year, doi_clean, all_of(country_codes)) %>%
  pivot_longer(cols = all_of(country_codes),
               names_to = "country_code", values_to = "value") %>%
  filter(!is.na(value), value > 0) %>%
  left_join(country_dict_std, by = c("country_code" = "code")) %>%
  mutate(country_name = coalesce(name, country_code)) %>%
  select(publication_year, doi_clean, country_code, country_name, value)



```



```{r}
summary_full <- df_long %>%
  group_by(publication_year, country_code, country_name) %>%
  summarise(count = n_distinct(doi_clean), .groups = "drop") %>%
  group_by(publication_year) %>%
  mutate(year_total = sum(count)) %>%
  ungroup()

all_countries <- sort(unique(summary_full$country_name))

```

simple addition
```{r}
countries_doi_count <- df_long  %>%
  group_by(country_name) %>%
  summarise(unique_items = n_distinct(doi_clean), .groups = 'drop')


```



```{r}
df_long |>
  summarise(
    rows         = n(),
    unique_country = n_distinct(country_name),
    unique_doi = n_distinct(doi_clean),
    # Year      = publication_year,
    .groups      = "drop"
  )
```

```{r}
# subset if unique items is greater than 10
more_common_countries <- countries_doi_count %>% 
  filter(unique_items >= 10)

```

analysis using only the top 100 countries
```{r}
more_common_countries_list <- more_common_countries$country_name

test_countries <- df_long %>% filter(country_name %in% more_common_countries_list)


test_countries |>
  summarise(
    rows         = n(),
    unique_country = n_distinct(country_name),
    unique_doi = n_distinct(doi_clean),
    # Year      = publication_year,
    .groups      = "drop"
  )
```

```{r}
library(gtsummary)
library(gt)

most_active_countries <- countries_doi_count %>% 
  filter(unique_items >= 1000)
# Create a publication-ready table 1

most_active_countries_table <- most_active_countries|> 
  arrange(desc(unique_items)) |> 
  gt() |> 
  cols_label(
    country_name = "Country of Author(s)",
    unique_items  = "Unique DOI's",
    # wt  = "Weight (1000 lbs)"
  ) |> 
  tab_header(
    title = "Top 10 Most Active Countries of Authorship",
    subtitle = "Base on OpenAlex"
  ) 

# my_table <- mtcars |> 
#   select(mpg, hp, wt) |> 
#   head() |> 
#   arrange(hp) |>  # Sorter by hp (lowest to highest)
#   gt() |> 
#   cols_label(
#     mpg = "Miles Per Gallon",
#     hp  = "Horsepower",
#     wt  = "Weight (1000 lbs)"
#   )

most_active_countries_table
```

```{r}
gtsave(most_active_countries_table, filename = "visual_output/most_active_countries_table.docx")

```


```{r}
most_active_countries_list <- most_active_countries$country_name

test1_countries <- df_long %>% filter(country_name %in% most_active_countries_list)


test1_countries |>
  summarise(
    rows         = n(),
    unique_country = n_distinct(country_name),
    unique_doi = n_distinct(doi_clean),
    # Year      = publication_year,
    .groups      = "drop"
  )
```

### journal hierarchy ###
downloaded from https://www.scimagojr.com/journalrank.php
first downloaded april 23, 2026
```{r}
#### SECTION with JOURNAL PRESTIGE ####
# read in Scimagojr data
library(readr)
scimagojr <- read_delim("scimagojr_data/scimagojr 2025.csv",
                             delim = ";", escape_double = FALSE, trim_ws = TRUE)
scimagojr <- read_delim(
  "scimagojr_data/scimagojr 2025.csv",
  delim          = ";",
  locale         = locale(decimal_mark = ","),
  show_col_types = FALSE
)

# View(scimagojr)

# subset scimagojr
scimagojr_trim <- scimagojr %>% 
  select(Title, Region, Categories, Areas, Overton, Issn,
         `H index`, `SJR Best Quartile`, `Citations / Doc. (2years)`,
         `Total Refs.`, `Total Citations (3years)`, `Total Docs. (2025)`, 
         `Total Docs. (3years)`, Rank
  )



```


```{r}
library(stringr)

normalize <- function(x) {
  x %>%
    str_to_lower() %>%
    str_remove_all("^the\\s+") %>%
    str_remove_all("[^a-z0-9 ]") %>%
    str_squish()
}

alex_clean    <- alex_doi_new1 %>% mutate(title_key = normalize(primary_location.source.display_name))
scimago_clean <- scimagojr_trim    %>% mutate(title_key = normalize(Title))

alex_doi_new_journal <- alex_clean %>%
  left_join(scimago_clean, by = "title_key")

# --- Pivot ISSN column into long format (one ISSN per row) ---
scimagojr_issn <- scimagojr %>%
  mutate(Issn = str_remove_all(Issn, "\\s")) %>%    # strip whitespace
  separate_rows(Issn, sep = ",") %>%                 # one row per ISSN
  filter(Issn != "", !is.na(Issn)) %>%               # drop blanks
  mutate(Issn = str_squish(Issn))

# --- Join on ISSN ---
# OpenAlex typically stores ISSNs as "XXXX-XXXX" so normalise both sides
normalize_issn <- function(x) str_remove_all(str_to_upper(x), "-")

alex_doi_new_journal <- alex_doi_new %>%
  mutate(issn_key = normalize_issn(primary_location.source.issn_l)) %>%
  left_join(
    scimagojr_issn %>% mutate(issn_key = normalize_issn(Issn)),
    by = "issn_key"
  )

# write csv
# write.csv(alex_doi_new_journal, file = "output_data/alex_doi_new_journal.csv")

# # table with all the journals and counts 
# n_journal <- n_distinct(alex_doi_new_journal$primary_location.source.display_name)

journal_counts <- alex_doi_new_journal %>%
  distinct(doi_clean, primary_location.source.display_name) %>% 
  count(primary_location.source.display_name, name = "Count") %>%
  select(primary_location.source.display_name, Count) %>% 
  rename("Journal Title" = primary_location.source.display_name) %>% 
  arrange(desc(Count))

journal_counts


```

Question to Claude Sonnet 5 - when i have a data frame ... I want to show a tree map using SJR best Quartile and Domain

original code 
library(dplyr)
library(treemapify)
library(ggplot2)

# 1. Join domain back in from alex_doi_new1 (which has it, post-merge)
alex_doi_new_journal <- alex_doi_new_journal %>%
  left_join(
    alex_doi_new1 %>% distinct(doi_clean, domain),
    by = "doi_clean"
  )# 2. Summarize: count of works by Domain x SJR Best Quartile
treemap_data <- alex_doi_new_journal %>%
  filter(!is.na(domain)) %>%
  mutate(
    sjr_quartile = if_else(is.na(`SJR Best Quartile`) | `SJR Best Quartile` == "",
                            "Unranked", `SJR Best Quartile`)
  ) %>%
  group_by(domain, sjr_quartile) %>%
  summarise(n_docs = n_distinct(doi_clean), .groups = "drop")

# 3. Treemap: rectangle size = doc count, grouped/bordered by Domain, colored by quartile
ggplot(treemap_data, aes(area = n_docs, fill = sjr_quartile,
                          label = sjr_quartile, subgroup = domain)) +
  geom_treemap() +
  geom_treemap_subgroup_border(colour = "white", size = 3) +
  geom_treemap_subgroup_text(place = "topleft", grow = FALSE,
                              alpha = 0.7, colour = "black",
                              fontface = "italic", size = 11) +
  geom_treemap_text(colour = "white", place = "center",
                     grow = FALSE, size = 10, reflow = TRUE) +
  scale_fill_manual(values = c(
    "Q1" = "#2c7fb8", "Q2" = "#41b6c4", "Q3" = "#a1dab4",
    "Q4" = "#ffffcc", "Unranked" = "grey70"
  )) +
  labs(title = "Documents by Domain and SJR Best Quartile",
       fill = "SJR Quartile") +
  theme(legend.position = "bottom")


Modified code
```{r}
library(dplyr)
library(treemapify)
library(ggplot2)

# 2. Summarize: count of works by Domain x SJR Best Quartile
treemap_data <- alex_doi_new_journal %>%
  filter(!is.na(Domain)) %>%
  mutate(
    sjr_quartile = if_else(is.na(`SJR Best Quartile`) | `SJR Best Quartile` == "",
                            "Unranked", `SJR Best Quartile`)
  ) %>%
  group_by(Domain, sjr_quartile) %>%
  summarise(n_docs = n_distinct(doi_clean), .groups = "drop")

# 3. Treemap: rectangle size = doc count, grouped/bordered by Domain, colored by quartile
ggplot(treemap_data, aes(area = n_docs, fill = sjr_quartile,
                          label = sjr_quartile, subgroup = Domain)) +
  geom_treemap() +
  geom_treemap_subgroup_border(colour = "darkgrey", size = 3) +
  geom_treemap_subgroup_text(place = "topleft", grow = FALSE,
                              alpha = 0.7, colour = "black",
                              fontface = "italic", size = 11) +
  geom_treemap_text(colour = "white", place = "center",
                     grow = FALSE, size = 10, reflow = TRUE) +
  scale_fill_manual(values = c(
    "Q1" = "#2c7fb8", "Q2" = "#41b6c4", "Q3" = "#a1dab4",
    "Q4" = "#ffffcc", "Unranked" = "grey70"
  )) +
  labs(title = "Documents by Domain and SJR Best Quartile",
       fill = "SJR Quartile") +
  theme(legend.position = "bottom")
```
What is a tabular way to represent this data?
response 
library(janitor)

alex_doi_new_journal %>%
  filter(!is.na(domain)) %>%
  mutate(sjr_quartile = if_else(is.na(`SJR Best Quartile`) | `SJR Best Quartile` == "",
                                 "Unranked", `SJR Best Quartile`)) %>%
  distinct(doi_clean, domain, sjr_quartile) %>%
  tabyl(domain, sjr_quartile) %>%
  adorn_totals(c("row", "col")) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns()   # shows "45.2% (120)" — percentage with raw count in parens
  
```{r}
library(janitor)

alex_doi_new_journal %>%
  filter(!is.na(Domain)) %>%
  mutate(sjr_quartile = if_else(is.na(`SJR Best Quartile`) | `SJR Best Quartile` == "",
                                 "Unranked", `SJR Best Quartile`)) %>%
  distinct(doi_clean, Domain, sjr_quartile) %>%
  tabyl(Domain, sjr_quartile) %>%
  adorn_totals(c("row", "col")) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns()   # shows "45.2% (120)" — percentage with raw count in parens
```



```{r}
# interactive graph
library(dplyr)
library(tidyr)

journal_year_counts <- alex_doi_new_journal %>%
  filter(!is.na(primary_location.source.display_name),
         !is.na(publication_year)) %>%
  distinct(
    doi_clean,
    primary_location.source.display_name,
    publication_year,
    `SJR Best Quartile`
    # `H index`
    # domain
  ) %>%
  count(
    primary_location.source.display_name,
    publication_year,
    `SJR Best Quartile`,
    # `H index`,
    # domain,
    name = "Citations"
  ) %>%
  rename(
    `Journal Title` = primary_location.source.display_name,
    Year = publication_year
  ) %>%
  arrange(`Journal Title`, Year)
```


Count only citations with ISSN
```{r}
alex_doi_issn <- alex_doi_new_journal |> 
  filter(!is.na(primary_location.source.issn_l)) #12555 out of 16111 citations had issn's

```


```{r}

```


```{r}
library(DT)

datatable(
  journal_year_counts,
  extensions = c("Buttons"),
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    order = list(list(0, "asc"), list(1, "asc"))
  ),
  filter = "top",
  rownames = FALSE
)

# # # left join title to journal title
# alex_doi_new_journal <- alex_doi_new %>%
#   left_join(scimagojr, by = c("primary_location.source.display_name" = "Title"))

# See how many failed to match
alex_doi_new_journal %>% 
  filter(is.na(`SJR Best Quartile`)) %>% 
  distinct(primary_location.source.display_name) %>% 
  nrow()

# unmatched journals
unmatched_journals <- alex_doi_new_journal %>% 
  filter(is.na(`SJR Best Quartile`)) %>% 
  distinct(primary_location.source.display_name) 
```

